---
title: "Assignment 3"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE}
source("code_execution.R")
```

# Introduction

The incarceration data gives valuable insight into the extreme exemplification
of racism that is the US prison system. I first calculated the mean proportion
of black inmates in the US prison system in 2018, which was calculated as
`r prop_black_inmates_2018`. This proportion is disturbingly high given the
actual percentage of black people within the total US population. The state with
the highest proportion of black inmates is `r state_highest_black_prop_2018`,
or Mississippi, with a proportion of `r highest_black_prop_overall`, which is
not exactly surprising as the South has existing legislature to hurt minorities.
The significant inequality increases throughout the years, with the year with
the highest proportion being `r highest_black_prop_overall_year`. In NY County,
the number of black inmates is incredibly disturbing, at a calculated total of
`r highest_number_of_black_inmates_2018` people.


```{r message=FALSE}
source("code_execution.R")
change_in_percentage_by_year
```

## Basic Plot

I included this plot to investigate the change in racial proportions of inmates 
over time. From 1991 to 2018, the proportion of white inmates saw a steady
decline, whereas that of black inmates saw a steady increase. This is likely due
to crime bills passed in 1986 and 1994 (both crafted by our current President).
The blatant disparity in drug related crimes, used to target black communities
and other oppressed minorites, is shown clearly in the data. Even with recent
"significant" reform to our racially rooted punishments, there had been little
adjustment in the percentages.

```{r message=FALSE}
source("code_execution.R")
private_vs_public_bar_chart
```

## Bar Chart

I included this bar chart to investigate for-profit, private prisons in the US
and their relation to racial manipulation. The relation is very clear, as
private prisons consistently show a higher proportion of black inmates. This
shows convincing evidence of modernized slavery in the US prison system. All
years included show a significantly higher proportion, indicating that black
people in America are more likely to be incarcerated for gains in underpaid
labor.

```{r message=FALSE}
source("code_execution.R")
percentage_black_illinois_map
```

## Map

This map shows the proportion of black inmates across the state of Illinois.
The highest proportions exist in the areas surrounding Chicago and St. Louis.
This is consistent with intuition, as urbanity of areas is directly proportional
to racial disparity in education and resources, forcing a higher proportion
of an oppressed population into crime. Police departments in urban areas are
also much more likely to pursue oppressed races/populations.
